---
title: "Assessment"
author: '100707631'
date: "2023-12-21"
output:
  word_document: default
  html_document:
    theme: united
    highlight: tango
  pdf_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#### 1.4.2 Load Necessary Libraries

library(quantmod) 
library(tidyverse) 
library(zoo)
library(dunn.test)
library(mlbench)
library(caret)
library(caTools)
library(MASS)
library(skimr)
```

# 1. Impact of Corona Pandemic on Apple's Stock Prices

## Paired T-Test

### **Purpose**

To compared stock prices of Apple in two specific periods (pre-pandemic
vs. during-pandemic) to see if there was a significant difference in the
mean closing prices between these two specific time frames. For that I
took pre-pandemic period (Jan 2019 to Dec 2019) and during pandemic
period (Jan 2020 to Dec 2020). The test compared each day from 2019 to
each day from 2020. First I conducted Paired T-test (Parametric test)
and after that I ran Wilcoxon Singned-Rank Test (non-parametric test) to
compare the results.

Null Hypothesis: There is no difference in the mean stock prices between
the two periods

**Pre-Pandemic Period**

```{r}
# Set the stock symbol, start date, and end date for the analysis
stock_symbol= "AAPL"
start_date4 <- "2019-01-01"
end_date4 <- "2019-12-31"
# Fetch historical stock price data from Yahoo Finance using quantmod
getSymbols(stock_symbol, src = "yahoo", from = start_date4, to = end_date4)
pre_pandemic <- Cl(get(stock_symbol))
print(pre_pandemic)

```

**Pandemic Period**

```{r}
# Set the stock symbol, start date, and end date for the analysis
start_date5 <- "2020-01-01"
end_date5 <- "2020-12-31"
# Fetch historical stock price data from Yahoo Finance using quantmod
getSymbols(stock_symbol, src = "yahoo", from = start_date5, to = end_date5)
During_pandemic <- Cl(get(stock_symbol))
```

### Cleaning Data

```{r}
print(length(pre_pandemic))
print(length(During_pandemic))

# Ensure equal length
min_length <- min(length(pre_pandemic), length(During_pandemic))
pre_pandemic_adjusted <- head(pre_pandemic, min_length)
During_pandemic_adjusted <- head(During_pandemic, min_length)

print(length(pre_pandemic_adjusted))
print(length(During_pandemic_adjusted))

sum(is.na(pre_pandemic_adjusted))
sum(is.na(During_pandemic_adjusted))

str(pre_pandemic_adjusted)
str(During_pandemic_adjusted)

# Convert 'xts' objects to numeric vectors
pre_pandemic_numeric <- as.numeric(coredata(pre_pandemic_adjusted$AAPL.Close))
During_pandemic_numeric <- as.numeric(coredata(During_pandemic_adjusted$AAPL.Close))
```

### **Checking Normality of Data**

```{r}
# Calculate the differences
differences <- pre_pandemic_numeric - During_pandemic_numeric
```

### **Performing Visual Inspection**

H0: Data is normally distributed

```{r}
# Histogram for visual inspection
hist(differences, main="Histogram of Differences", xlab="Differences", col="blue")


```

```{r}

# Q-Q Plot for visual inspection
qqnorm(differences)
qqline(differences, col="red")
```

```{r}

# Shapiro-Wilk normality test
shapiro.test(differences)
```


The null hypothesis is not true as p-value is lower than the set p
value. Which means the data is not normally distributed.

### Performing Paired T-Test

```{r}

# Perform the paired t-test on these numeric vectors
t.test_result <- t.test(pre_pandemic_numeric, During_pandemic_numeric, paired = TRUE)

# Print the result of the t-test
print(t.test_result)


```

**Analysis**

Paired T-test compares the mean stock prices between two periods,
pre-pandemic and during the pandemic, for Apple stock. The paired t-test
is appropriate here because it compares the same stock across two
different time periods, thus accounting for the pairing of data points.
The test results rejected the null hypothesis.

Here's a breakdown of the result:

-   **t-value (-42.612):** This is a measure of the difference between
    the two periods relative to the variability in the sample data. A
    very high or very low t-value (far from zero) indicates a
    significant difference. In my case, -42.612 is a large negative
    value, indicating a significant decrease in the stock prices during
    the pandemic compared to the pre-pandemic period.

-   **Degrees of freedom (df = 250):** This is calculated based on the
    number of paired observations in my data. It's used to determine the
    critical value of t at a given significance level.

-   **p-value (\< 2.2e-16):** This is extremely small, much lower than
    typical significance levels (like 0.05 or 0.01), suggesting that the
    observed difference in means is highly unlikely to have occurred by
    random chance. In other words, there's a statistically significant
    difference in the stock prices between the two periods.

-   **95% Confidence Interval (-45.05698 to -41.07599):** This interval
    gives a range of values that are likely to contain the true mean
    difference between the two periods. The fact that this interval does
    not include 0 strengthens the conclusion that there's a significant
    difference.

-   **Mean difference (-43.06648):** This is the average difference in
    stock prices between the pre-pandemic and during pandemic periods.
    Since it's negative, it suggests that the stock price was, on
    average, lower during the pandemic than before it.

### Visualization

```{r}
plot(pre_pandemic_numeric, type='l', col='blue', main='AAPL Stock Prices: Pre vs. During Pandemic', xlab='Date', ylab='Closing Price', ylim = c(0,150))
lines(During_pandemic_numeric, col='red')
legend("topright", legend=c("Pre-Pandemic", "During Pandemic"), col=c("blue", "red"), lty=1)

```

#### **Graph Analysis**

This graph represents the closing prices of AAPL (Apple Inc.) stock
before and during the pandemic, plotted over time. Here's the breakdown
of the graph:

-   **X-axis:** The x-axis is labeled as 'Date'. It represents time,
    although specific dates are not given. The data points likely
    correspond to trading days, with the index starting at 0 and going
    up to around 250, which would roughly correspond to the number of
    trading days in a year (minus weekends and holidays).

-   **Y-axis:** The y-axis is labeled as 'Closing Price' and ranges from
    0 to 150. This axis represents the closing price of AAPL stock for
    each day in the dataset.

-   **Blue Line (Pre-Pandemic):** This line represents the closing
    prices of AAPL stock during the pre-pandemic period. The data points
    are connected by a line (indicated by type='l' in the code), and the
    line is colored blue.

-   **Red Line (During Pandemic):** This line represents the closing
    prices of AAPL stock during the pandemic period. It is overlaid on
    the same graph as the pre-pandemic data and is colored red.

Analyzing the graph, we can observe the following:

-   The blue line (pre-pandemic) starts off lower and shows some
    fluctuation over time but tends to stay below a certain level
    throughout the period.

-   The red line (during the pandemic) starts at a higher level than
    where the blue line left off, indicating an increase in the stock
    price at the beginning of the pandemic period compared to the end of
    the pre-pandemic period.

-   The red line shows more volatility and overall trends upwards,
    reaching higher closing prices than the pre-pandemic period. There
    appears to be a general upward trend in the stock price during the
    pandemic, which could be attributed to various factors such as
    changes in market conditions, investor sentiment, and company
    performance, among others.

### 2.4 Wilcoxon Singned-Rank Test

Null Hypothesis: There is no difference in the median stock prices
between the two periods

Here's how the Wilcoxon signed-rank test worked:

-   **Pairing of Data:** Each observation from one period is paired with
    the corresponding observation from the other period. In the context
    of stock prices, each day's closing price in the pre-pandemic period
    is paired with the closing price on the same day during the
    pandemic.

-   **Ranking of Differences:** For each pair, calculated the difference
    and then rank these differences, ignoring the sign. The smallest
    non-zero difference gets rank 1, the second smallest rank 2, and so
    on.

-   **Sum of Ranks:** Sum the ranks for the differences that were
    positive and sum the ranks for the differences that were negative.

-   **Test Statistic (V):** The test statistic for the Wilcoxon
    signed-rank test is the smaller of the two sums of ranks. If there's
    a large difference between these sums, it suggests a difference
    between the two periods.

-   **Significance:** The p-value is calculated from the test statistic
    to determine the significance of the result. If the p-value is less
    than the chosen alpha level (0.05), then the null hypothesis of no
    difference is rejected.

```{r}
# Perform the Wilcoxon signed-rank test
wilcox_test_result <- wilcox.test(pre_pandemic_numeric, During_pandemic_numeric, paired = TRUE)

print(wilcox_test_result)

```

#### Analyzing the Wilcoxon Test Output:

-   **Test Statistic (V = 0):** This suggests that all the ranks for the
    differences pointed in the same direction (i.e., either all positive
    or all negative). This indicates a consistent difference between the
    pre-pandemic and during-pandemic prices.

-   **p-value (\< 2.2e-16):** This extremely low p-value indicates that
    the observed result is highly unlikely under the null hypothesis.
    Therefore, we reject the null hypothesis that there is no difference
    in the mean ranks of stock prices between the two periods.

-   **Alternative Hypothesis:** The alternative hypothesis stated by the
    test is that there is a true location shift (difference in medians)
    between the pre-pandemic and during-pandemic periods. The test
    result supports this alternative hypothesis.

#### 2.4.1 Comparison between paired test result and Wilcoxon test result

Both tests give a very significant result (p-value \< 2.2e-16),
indicating that there is a statistically significant difference between
the two time periods.

-   Despite the violation of assumptions for the paired t-test, the
    p-value is extremely low for both tests, which can happen with large
    sample sizes or with very pronounced effects. The Central Limit
    Theorem might be at play here, helping the t-test to still give a
    robust result despite the assumption violation.

-   The paired t-test provides an estimate of the mean difference and a
    confidence interval, which are not available from the Wilcoxon test.
    The Wilcoxon test, on the other hand, indicates a significant shift
    in the median value without assuming normality.

# 2. Comparison of stock prices using ANOVA

**Purpose**

Performing ANOVA test to compare the average closing prices across three
distinct periods (pre-pandemic, during-pandemic, post-pandemic).

H0: equal average closing prices across the three time periods

This is how ANOVA worked:

-   Grouping: First, we need to have independent samples grouped by a
    factor. In my case, the factor is 'TimePeriod' with three levels:
    pre-pandemic, during-pandemic, and post-pandemic.

-   Assumptions: ANOVA assumes that the data within and across groups
    are normally distributed and that the variances are equal
    (homoscedasticity). It also assumes that the observations are
    independent.

-   Null Hypothesis (H0): The null hypothesis in ANOVA states that there
    are no differences between the group means, and any observed
    differences are due to random variation.

-   Alternative Hypothesis (Ha): The alternative hypothesis is that at
    least one group mean is different from the others.

### Fetching Post Pandemic Stock Prices

```{r}
stock_symbol <- "AAPL"
start_date_p <- "2021-01-01"
end_date_p <- "2022-12-31"


getSymbols(stock_symbol, src = "yahoo", from = start_date_p, to = end_date_p)
post_pandemic <- Cl(get(stock_symbol))


```

### Extract the closing prices and create a vector for time periods

```{r}
pre_pandemic_prices <- coredata(pre_pandemic)
during_pandemic_prices <- coredata(During_pandemic)
post_pandemic_prices <- coredata(post_pandemic)

pre_pandemic_period <- rep("Pre-Pandemic", length(pre_pandemic_prices))
during_pandemic_period <- rep("During-Pandemic", length(during_pandemic_prices))
post_pandemic_period <- rep("Post-Pandemic", length(post_pandemic_prices))

```

### Combine the data into a single dataframe

```{r}

combined_data <- data.frame(
  ClosingPrice = c(pre_pandemic_prices, during_pandemic_prices, post_pandemic_prices),
  TimePeriod = factor(c(pre_pandemic_period, during_pandemic_period, post_pandemic_period))
)


```

### Performing ANOVA

```{r}

anova_result <- aov(ClosingPrice ~ TimePeriod, data = combined_data)
summary(anova_result)
```

**ANOVA Output Interpretation**

The very small p-value (less than 0.0000000000000002) means that there
is strong statistical evidence to reject the null hypothesis of equal
average closing prices across the three time periods (pre-pandemic,
during-pandemic, post-pandemic). In other words, the average closing
price of AAPL is significantly different across at least two of the
compared time periods.

### Visual Respresentation (Box plot)

```{r}
ggplot(combined_data, aes(x = TimePeriod, y = ClosingPrice)) +
  geom_boxplot() +
  labs(title = "Box Plot of AAPL Closing Prices by Time Period",
       x = "Time Period",
       y = "Closing Price")
```

### TUKEY HSD Test

Tukey's test compares the means of every group to the means of every
other group. This is important after an ANOVA because, while ANOVA can
tell you that there is a significant difference somewhere among the
groups, it doesn't specify where that difference lies.

```{r}
# Perform Tukey's HSD test
tukey_test <- TukeyHSD(anova_result)

# Print the results of the Tukey test
print(tukey_test)

```

**Pairwise Mean Differences Analysis**

1.  **Post-Pandemic vs. During-Pandemic**:

    -   Difference (**diff**): +52.69956

    -   The mean closing price during the post-pandemic period is higher
        than the during-pandemic period by approximately 52.70 units.

    -   The 95% confidence interval (**`lwr`** to **`upr`**) does not
        contain 0, which indicates a statistically significant
        difference.

    -   The p-value (**`p adj`**) is 0, meaning the difference is
        significant at the 95% confidence level.

2.  **Pre-Pandemic vs. During-Pandemic**:

    -   Difference (**`diff`**): -43.21995

    -   The mean closing price during the pre-pandemic period is lower
        than the during-pandemic period by approximately 43.22 units.

    -   The 95% confidence interval does not contain 0, indicating a
        statistically significant difference.

    -   The p-value is 0, signifying the difference is significant at
        the 95% confidence level.

3.  **Pre-Pandemic vs. Post-Pandemic**:

    -   Difference (**`diff`**): -95.91952

    -   The mean closing price during the pre-pandemic period is lower
        than the post-pandemic period by approximately 95.92 units.

    -   The 95% confidence interval does not contain 0, indicating this
        difference is statistically significant.

    -   The p-value is 0, meaning the difference is significant at the
        95% confidence level.

**Interpretation**

All pairwise comparisons are statistically significant, indicating that
the average closing prices of Apple's stock are significantly different
in all three periods:

-   The stock price increased significantly from the during-pandemic
    period to the post-pandemic period.

-   The stock price decreased significantly from the pre-pandemic period
    to the during-pandemic period.

-   There is a significant increase in the stock price from the
    pre-pandemic to the post-pandemic period.

### **Conclusions**

These results suggest that the pandemic had a noticeable impact on the
stock price of Apple, with prices being significantly lower during the
pandemic compared to both before and after. The post-pandemic recovery
also indicates a significant rise in stock prices compared to
pre-pandemic levels

## Kruskal-Wallis H test

To apply the Kruskal-Wallis H test, which is the non-parametric
alternative to ANOVA, it will need to have the data organized by groups,
just as it would for ANOVA.The Kruskal-Wallis H test will test the null
hypothesis that the median of all groups is equal, against the
alternative hypothesis that at least one group has a different median.

H0: The median of all groups is equal

H1: At least one group has a different median.

```{r}
kruskal.test_result <- kruskal.test(ClosingPrice ~ TimePeriod, data = combined_data)

# Print the result of the Kruskal-Wallis H test
print(kruskal.test_result)
```

**Output Interpretation**

Given that the p-value is extremely small, we can reject the null
hypothesis of the Kruskal-Wallis test, which states that the medians of
all groups are equal. This means that the test has found strong evidence
to suggest that not all the groups have the same median closing price.

However, while the Kruskal-Wallis test tells us that there is a
significant difference, it does not tell us which specific groups are
different from each other. To identify the specific groups between which
there are differences, you would need to conduct post-hoc pairwise
comparisons. This is typically done with a Dunn's test, which compares
the medians of the groups in a pairwise manner while adjusting for the
multiple comparisons problem.

### Dunn Test

```{r}
# Run the Dunn test for post-hoc analysis after Kruskal-Wallis
dunn_test_result <- dunn.test(combined_data$ClosingPrice, combined_data$TimePeriod, method="bonferroni")

# Print the Dunn test result
print(dunn_test_result)
```

**Output Interpretation**

The output from the Dunn test provides the following information:

1.  **Chi-squared**: The Chi-squared statistic from the Kruskal-Wallis
    test is 827.3926, which is just a reference point already known from
    the initial test.

2.  **Z-values**: These are the standardized differences between the
    ranks of the groups. There are three Z-values provided, each
    corresponding to a pairwise comparison between the groups:

    -   During-Pandemic vs. Post-Pandemic: -16.567341

    -   During-Pandemic vs. Pre-Pandemic: 9.659462

    -   Post-Pandemic vs. Pre-Pandemic: 27.691807

    The negative Z-value indicates that the median of the first group
    (During-Pandemic) is lower than that of the second group
    (Post-Pandemic), while positive Z-values indicate the opposite.

3.  **P-values**: These are the p-values for each of the pairwise
    comparisons:

    -   During-Pandemic vs. Post-Pandemic: 6.00e-62

    -   During-Pandemic vs. Pre-Pandemic: 2.24e-22

    -   Post-Pandemic vs. Pre-Pandemic: 4.38e-169

    These p-values are very low, indicating that there are significant
    differences between each pair of groups.

4.  **Adjusted P-values**: These are the p-values adjusted for multiple
    comparisons using the Bonferroni method:

    -   During-Pandemic vs. Post-Pandemic: 1.80e-61

    -   During-Pandemic vs. Pre-Pandemic: 6.72e-22

    -   Post-Pandemic vs. Pre-Pandemic: 1.31e-168

    Even after adjustment for multiple comparisons, the p-values remain
    extremely low, indicating strong evidence against the null
    hypothesis in each pairwise comparison.

5.  **Comparisons**: This indicates which groups were compared to each
    other.

The results suggest an increase in stock prices from the pre-pandemic to
during-pandemic period and then an even more significant increase from
the during-pandemic to the post-pandemic period. These results matches
with what we got from ANOVA test.

# 3. Impact of the pandemic on Apple Inc.'s (AAPL) stock price and trading volume

**Purpose**

The primary objective of this analysis is to investigate the
relationship between the closing prices of Apple Inc. (AAPL) and its
trading volume. This study is predicated on the hypothesis that trading
volume, as a measure of market activity and investor interest, can have
a significant impact on the stock price. By examining this relationship,
the analysis aims to uncover insights into how market participation
influences the value of AAPL shares. Closing value is dependent variable
and trading volume is independent variable.

### Getting Trading Volume Data

```{r}
stock_symbol <- "AAPL"
start_date_t <- as.Date("2019-01-01")
end_date_t <- as.Date("2019-12-01")  

# Fetch the stock data
getSymbols(stock_symbol, src = "yahoo", from = start_date_t, to = end_date_t)
pre_aapl_volume <- Vo(get(stock_symbol))


stock_symbol <- "AAPL"
start_date_t1 <- as.Date("2020-01-01")
end_date_t1 <- as.Date("2020-12-01")  
getSymbols(stock_symbol, src = "yahoo", from = start_date_t1, to = end_date_t1)
during_aapl_volume <- Vo(get(stock_symbol))

stock_symbol <- "AAPL"
start_date_t2 <- as.Date("2021-01-01")
end_date_t2 <- as.Date("2022-12-01")  
getSymbols(stock_symbol, src = "yahoo", from = start_date_t2, to = end_date_t2)
post_aapl_volume <- Vo(get(stock_symbol))

```

### Combine all intervals in one dataframe

```{r}
# Combine data for each period into a dataframe
length(pre_pandemic_prices)
length(pre_aapl_volume)

length(during_pandemic_prices)
length(during_aapl_volume)

length(post_pandemic_prices)
length(post_aapl_volume)

# Trimming the data (assuming the extra data points are at the end)
pre_pandemic_prices <- pre_pandemic_prices[1:231]
during_pandemic_prices <- during_pandemic_prices[1:231]
post_pandemic_prices <- post_pandemic_prices[1:482]

# Now recreate the data frames
pre_pandemic_data <- data.frame(Price = pre_pandemic_prices, Volume = pre_aapl_volume, Period = "Pre-Pandemic")
during_pandemic_data <- data.frame(Price = during_pandemic_prices, Volume = during_aapl_volume, Period = "During-Pandemic")
post_pandemic_data <- data.frame(Price = post_pandemic_prices, Volume = post_aapl_volume, Period = "Post-Pandemic")

# Combine into one dataframe
combined_data <- rbind(pre_pandemic_data, during_pandemic_data, post_pandemic_data)

```

### **Visual Inspection**

```{r}
# Scatter plot for the pre-pandemic period
ggplot(pre_pandemic_data, aes(x = pre_aapl_volume, y = pre_pandemic_prices)) +
  geom_point(alpha = 0.5) +
  labs(title = "Scatter Plot of AAPL Price vs Volume (Pre-Pandemic)",
       x = "Volume",
       y = "Closing Price") +
  theme_minimal()


```

Here are some observations and analyses of the plot:

1.  **Volume on the X-axis:** The volume is plotted on the horizontal
    axis (X-axis) and is in the range from 0 to over 3 x 10\^8 (or 300
    million). The volume represents the number of shares traded during a
    single trading session.

2.  **Closing Price on the Y-axis:** The closing price is plotted on the
    vertical axis (Y-axis) and ranges approximately between \$40 and
    above \$60. The closing price is the last price at which the stock
    traded during a regular trading session on a given day.

3.  **Data Points:** Each dot on the scatter plot represents a single
    day's trading activity, with its position determined by that day's
    trading volume (x-coordinate) and the stock's closing price
    (y-coordinate).

4.  **Density of Points:** There appears to be a higher concentration of
    points with volumes in the lower range (closer to 0) and closing
    prices around the \$50 to \$60 range. This suggests that on most
    days, the trading volume was on the lower end of the spectrum, with
    closing prices frequently settling in the middle of the price range.

5.  **Outliers:** There are a few days with very high trading volumes
    that stand out from the rest of the data points. It's not clear from
    the plot whether these higher volumes are associated with higher or
    lower closing prices, as the points are quite spread out.

6.  **Trend Analysis:** The plot does not show a clear linear
    relationship between volume and closing price. There's quite a bit
    of scatter in the data points, indicating that the closing price of
    AAPL stock during the pre-pandemic period was not strongly or
    consistently associated with the trading volume.

7.  **Opacity (alpha = 0.5):** The points are plotted with some
    transparency, which is helpful for visualizing the density of the
    points and understanding where there are overlaps. Where you see
    darker areas, it indicates a higher number of trading days with
    similar volume and price characteristics.

8.  **Interpretation:** Without a clear trend, we can say that during
    the pre-pandemic period, trading volume was not a reliable predictor
    of the closing price for AAPL stock. Other factors not shown on this
    chart are likely at play in determining the closing price.

### **Performing Linear Regression**

```{r}

lm_model <- lm(Price ~ AAPL.Volume + Period, data = combined_data)
summary(lm_model)

```

**Regression Test Output Interpretation**

The negative coefficient for **`AAPL.Volume`** implies that an increase
in trading volume is associated with a decrease in the stock price,
although the effect size is small. This finding directly relates to the
hypothesis about the impact of market activity and investor interest (as
reflected by trading volume) on stock price. This result provides
insights into how market participation (measured via trading volume)
influences the value of AAPL shares. It suggests that higher trading
volumes might be associated with periods of stock price decline, which
could be due to various factors like market sell-offs, increased
investor activity during price dips, etc.\
The inclusion of the **`Period`** variable in regression model allows to
assess how stock prices varied during different phases of the pandemic
(pre, during, and post). The significant coefficients for the
pre-pandemic and post-pandemic periods compared to the baseline
(during-pandemic) show how the stock price differed across these times.

# 4. Post Pandemic Comparison Between Apple and Microsoft Stocks

**Problem Statement**

A comparative analysis of AAPL and MSFT stock prices could provide
valuable insights into their relative market performance over a
specified time period.

**Null Hypothesis**

(HO): There is no significant difference in the average stock prices of
Apple Inc. (AAPL) and Microsoft Corporation (MSFT) during the specified
time period.

This hypothesis sets the basis for the unpaired t-test, which will be
used to statistically determine if the observed differences in average
stock prices between AAPL and MSFT are significant or could have
occurred by random chance. Rejecting this null hypothesis would imply
that there is a statistically significant difference in the average
stock prices of these two companies, providing valuable insights into
their relative performance.

## Unpaired T-Test

### Fetch Microsoft Data from Yahoo Finance

```{r}
msft_symbol <- "MSFT"
aapl_symbol <- "AAPL"

start_date <- "2022-01-01" 
end_date <- "2023-12-26"

# Fetch AAPL data
getSymbols(aapl_symbol, src = "yahoo", from = start_date, to = end_date)
aapl_data <- Cl(get(aapl_symbol))
print(aapl_data)

# Fetch MSFT data
getSymbols(msft_symbol, src = "yahoo", from = start_date, to = end_date)
msft_data <- Cl(get(msft_symbol))
print(msft_data)
```

### Prepare Data for the T-Test

```{r}
# Find the overlapping date range
common_start_date <- max(index(aapl_data)[1], index(msft_data)[1])
common_end_date <- min(tail(index(aapl_data), n=1), tail(index(msft_data), n=1))

# Print the common date range
print(paste("Common date range: ", common_start_date, " to ", common_end_date))

# Truncate AAPL data to the common date range
aapl_data_common <- window(aapl_data, start=common_start_date, end=common_end_date)

# Truncate MSFT data to the common date range
msft_data_common <- window(msft_data, start=common_start_date, end=common_end_date)

```

### Ensuring Exactly the Same Time Frame for Both AAPL and MSFT

```{r}
aapl_data_common <- window(aapl_data, start="2022-01-03", end="2023-12-22")
msft_data_common <- window(msft_data, start="2022-01-03", end="2023-12-22")

```

### Running Unpaired Tests

```{r}

t_test_result <- t.test(aapl_data_common[,1], msft_data_common[,1])

# Print the results
print(t_test_result)

```

**Unpaired T-test Output Interpretation:**

1.  **t-Statistic (t)**: -64.395

    -   This is a measure of the difference between the two sample means
        relative to the variability in data. The negative value
        indicates that the mean stock price of AAPL is lower than that
        of MSFT for the period analyzed.

2.  **Degrees of Freedom (df)**: 677.21

    -   This number, which can be a non-integer value in the Welch
        t-test, indicates the degrees of freedom for the test. It's used
        in determining the p-value.

3.  **P-Value**: \< 2.2e-16

    -   The p-value is extremely small, far below the common
        significance level of 0.05. This suggests that the observed
        difference in the average stock prices between AAPL and MSFT
        during this period is statistically significant.

4.  **Alternative Hypothesis**: True difference in means is not equal to
    0

    -   The test results strongly support the alternative hypothesis,
        indicating that there is a significant difference in the average
        stock prices of AAPL and MSFT.

5.  **95 Percent Confidence Interval**: [-131.1592, -123.3975]

    -   This interval suggests that the true difference in the mean
        stock prices between AAPL and MSFT lies somewhere in this range.
        Both ends of the interval are negative, reinforcing the finding
        of a higher average price for MSFT compared to AAPL.

6.  **Sample Estimates**: Mean of x (AAPL) = 163.4377, Mean of y (MSFT)
    = 290.7161

    -   The mean stock prices for AAPL and MSFT are provided, with MSFT
        having a notably higher average stock price during the period.

### Visualization

```{r}

#Converting xts objects into numerical data for boxplot
aapl_prices <- as.numeric(coredata(aapl_data_common))
msft_prices <- as.numeric(coredata(msft_data_common))


# Creating a boxplot for AAPL and MSFT stock prices
boxplot(aapl_prices, msft_prices, names = c("AAPL", "MSFT"), main="AAPL vs. MSFT Stock Prices", ylab="Closing Price")


```

The box plot visualizes the differences in stock prices between AAPL and
MSFT and indicates that MSFT had higher median stock prices with a wider
range of values. The fact that there's no overlap between the boxes of
AAPL and MSFT suggests that the stock prices for these two companies
were quite distinct, with MSFT consistently higher during this period.

### Conclusion

The significant difference in means, as evidenced by the unpaired
t-test, indicates that MSFT's stock price was, on average, higher than
AAPL's during the specified period. The wide gap in the confidence
interval and the large magnitude of the t-statistic underscore the
robustness of this difference.

The results provide a clear statistical backing to the observation that
Microsoft's stock has outperformed Apple's in terms of average stock
price over the selected time frame.

## Two-Sample Wilcoxon Rank Sum Test (Mann-Whitney)

Since I've already found a significant difference with the Welch Two
Sample t-test (a version of the t-test that does not assume equal
variances), it will be interesting to see if the Mann-Whitney U test
aligns with this finding despite not requiring the normality assumption.

```{r}
# Summary statistics for AAPL and MSFT data
#summary(aapl_data)
#summary(msft_data)

# Check data types
str(aapl_data)
str(msft_data)

# Check the range of dates for AAPL and MSFT
range(index(aapl_data))
range(index(msft_data))

# Check if all dates in AAPL data are in MSFT data and vice versa
all(index(aapl_data) %in% index(msft_data))
all(index(msft_data) %in% index(aapl_data))

# Find the overlapping date range
common_start_date <- max(index(aapl_data)[1], index(msft_data)[1])
common_end_date <- min(end(index(aapl_data)), end(index(msft_data)))

# Print the common date range
print(paste("Common date range: ", common_start_date, " to ", common_end_date))


# Boxplot for visual inspection
boxplot(as.numeric(aapl_data_common[,1]), as.numeric(msft_data_common[,1]),
        names=c("AAPL", "MSFT"), main="Boxplot of Closing Prices")


# Data has been cleaned and verified
mw_test_result <- wilcox.test(aapl_prices, msft_prices)
print(mw_test_result)

```

**Mann-Whitney Test's Output Interpretation**

Given the p-value is extremely low, we can confidently reject the null
hypothesis (H0) that there is no significant difference in the median
stock prices between Apple Inc. (AAPL) and Microsoft Corporation (MSFT)
during the specified time period.

However, the W statistic of 0 is unusual and may indicate an extreme
result. Typically, such a result could occur if one of the distributions
is entirely above or below the other, leading to a situation where one
group consistently has higher or lower values than the other group
without any overlap. To be sure the results are right, we verified data
integrity before conducting Wilcoxon Rank Sum Test.

# 5. Market Movement Synchronization

**Objective**: To understand the degree of synchronization between Apple
and Microsoft's stock movements over the past two years using Pearson
Correlation test.

## Pearson Correlation Test

```{r}

correlation_coefficient <- cor(aapl_data_common, msft_data_common, method = "pearson")
print(correlation_coefficient)

```

#### Pearson Correlation Test Output Interpretation:

1.  **Strong Positive Correlation**: A coefficient of 0.937 indicates
    that the movements in Apple's and Microsoft's stock prices are very
    closely aligned. Generally, as the price of one goes up, the price
    of the other tends to go up as well, and vice versa.

2.  **Market and Sector Influences**: This high correlation might
    suggest that both stocks are influenced by similar factors, such as
    market trends, economic conditions, or developments in the
    technology sector.

### Visual Representation using Scatter Plot

```{r}
ggplot(data = data.frame(aapl_data_common, msft_data_common), aes(x = aapl_data_common, y = msft_data_common)) +
  geom_point() + 
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Scatter Plot of AAPL vs MSFT Closing Prices",
       x = "AAPL Closing Price",
       y = "MSFT Closing Price")
```

**Scatter Plot Analysis**

1.  **Data Points Distribution**: The data points are densely packed and
    form a clear upward trend from the bottom left to the top right.
    This indicates a positive correlation between AAPL and MSFT closing
    prices, meaning they tend to move in the same direction.

2.  **Trend Line**: The blue line represents the best fit line through
    the data points, calculated using linear regression. It has a
    positive slope, which confirms the positive correlation between the
    two stocks' closing prices.

3.  **Strength of the Correlation**: The data points closely follow the
    trend line, which indicates a strong linear relationship. This
    visual supports the Pearson correlation coefficient of approximately
    0.937 that I calculated, demonstrating a strong positive
    correlation.

## Spearman Correlation Test

```{r}
# Calculate Spearman's Rank Correlation Coefficient
spearman_correlation_coefficient <- cor(aapl_data_common, msft_data_common, method = "spearman")

# Print the Spearman's Rank Correlation Coefficient
print(spearman_correlation_coefficient)

```

**Spearman Correlation Coefficient (0.9476658)**: This value is also
close to +1, indicating a very strong positive monotonic relationship
between the ranked values of AAPL and MSFT closing prices. The Spearman
correlation is less sensitive to outliers and does not require the
assumption of normally distributed data, so this high correlation
suggests that the relationship is strong even without the assumption of
linearity.

Both correlation coefficients (Pearson and Spearman) being high and very
similar suggests that the relationship is both linear and monotonic. In
practical terms, it implies that AAPL and MSFT stocks have moved
together quite closely during the period analyzed. This could be due to
various factors, such as the companies' performances, market trends, or
economic conditions affecting the technology sector in which both
companies operate.

# 6. Predicting the Breast Cancer Tumor Class

I took a new data set called Breast Cancer which has 11 columns and 699
rows. The purpose is to correctly identify the class of the tumor using
different exploratory variables.

When considering a logistic regression model for predicting whether a
tumor is benign or malignant based on the breast cancer dataset, we need
to identify potential explanatory (independent) variables that may
influence the dependent variable, which in this case is the "Class"
(malignant = 1, benign = 0).

Here's a breakdown of the columns of dataset and their potential as
explanatory variables:

-   Index/ID: This is likely a unique identifier for each observation
    and does not provide any predictive power for the model.

-   Cell Thickness (Cl.thickness): This could be an important
    explanatory variable as it relates to how thick the cell formations
    are in the breast tissue. Abnormal cell thickness can be indicative
    of cancerous changes.

-   Cell Size (Cell.size): The size of the cells could be a significant
    predictor as cancer cells often vary in size and can be larger or
    smaller than normal cells.

-   Cell Shape (Cell.shape): The shape of cells is another key
    indicator. Irregularly shaped cells are often associated with
    malignancy.

-   Marginal Adhesion (Marg.adhesion): This measures how cells stick to
    their neighboring cells. Changes in adhesion properties can be a
    sign of malignancy.

-   Epithelial Cell Size (Epith.c.size): Variations in the size of
    epithelial cells can be indicative of cancerous tissue.

-   Bare Nuclei (Bare.nuclei): This refers to the presence of nuclei
    without surrounding cytoplasm. It's a significant factor in many
    cancer diagnoses.

-   Bland Chromatin (Bl.cromatin): The texture of the chromatin in the
    nucleus can provide information about whether a tumor is benign or
    malignant.

-   Normal Nucleoli (Normal.nucleoli): Nucleoli within the cell nucleus
    can change in appearance in cancer cells.

-   Mitoses: The rate of cell division (mitosis) can be an indicator of
    cancer, as cancer cells often divide more rapidly than normal cells.


### Load Dataset

```{r}
# Load the dataset
data("BreastCancer")

```

### Descriptive Statistics

```{r}

BreastCancer_no_Id <- BreastCancer[, !names(BreastCancer) %in% c("Id")]

# Convert all columns to numeric, except 'Class'
BreastCancer_no_Id[setdiff(names(BreastCancer_no_Id), "Class")] <- 
  lapply(BreastCancer_no_Id[setdiff(names(BreastCancer_no_Id), "Class")], 
         function(x) as.numeric(as.character(x)))

# Convert 'Class' to binary (0 and 1)
BreastCancer_no_Id$Class <- ifelse(BreastCancer_no_Id$Class == "benign", 0, 1)

# Descriptive Statistics
summary_stats <- data.frame(
  Mean = apply(BreastCancer_no_Id, 2, function(x) mean(x, na.rm = TRUE)),
  Median = apply(BreastCancer_no_Id, 2, function(x) median(x, na.rm = TRUE)),
  SD = apply(BreastCancer_no_Id, 2, function(x) sd(x, na.rm = TRUE))
)

print(summary_stats)



```

### **Data Cleaning**

```{r}

# Summarize missing values in each column
na_count <- sapply(BreastCancer, function(x) sum(is.na(x)))
na_count

# Convert Bare.nuclei to numeric
BreastCancer$Bare.nuclei <- as.numeric(as.character(BreastCancer$Bare.nuclei))

# Check if conversion introduced any new NAs
na_count_bare_nuclei <- sum(is.na(BreastCancer$Bare.nuclei))
print(paste("Number of NAs in Bare.nuclei after conversion:", na_count_bare_nuclei))

# Impute missing values in Bare.nuclei with median
BreastCancer$Bare.nuclei[is.na(BreastCancer$Bare.nuclei)] <- median(BreastCancer$Bare.nuclei, na.rm = TRUE)

# Verify that there are no more NAs
na_count_updated <- sapply(BreastCancer, function(x) sum(is.na(x)))
na_count_updated

#Checking for unique identifiers 
sapply(BreastCancer, function(x) length(unique(x)))
sapply(BreastCancer, function(x) sum(is.na(x)))
```

### **Data Preparation**

```{r}

# Remove the Id column
BreastCancer <- BreastCancer[, !names(BreastCancer) %in% c("Id")]
View(BreastCancer)

# Data preparation
set.seed(123) # Setting seed for reproducibility


# Convert factors to numeric values if necessary
BreastCancer$Class <- ifelse(BreastCancer$Class == "malignant", 1, 0)

# Splitting the dataset into training and testing sets
split <- sample.split(BreastCancer$Class, SplitRatio = 0.7)
train_data <- subset(BreastCancer, split == TRUE)
test_data <- subset(BreastCancer, split == FALSE)

# Separating predictors and response
train_x <- train_data[, -ncol(train_data)]
train_y <- train_data$Class
test_x <- test_data[, -ncol(test_data)]
test_y <- test_data$Class

```

### **Fitting a Logistic Regression Model**

```{r}

# Fit logistic regression model
model <- glm(Class ~ ., data = train_data, family = binomial())

# Summarize the model
summary(model)

```

### **Evaluating the Model**

```{r}
# Making predictions
predictions <- predict(model, newdata = test_x, type = "response")
predicted_class <- ifelse(predictions > 0.5, 1, 0)

# Confusion Matrix
confusionMatrix <- table(test_y, predicted_class)
print(confusionMatrix)

# Calculate accuracy
accuracy <- sum(diag(confusionMatrix)) / sum(confusionMatrix)
print(paste("Accuracy:", accuracy))

```

Confusion Matrix:

**True Negatives (TN) - 132:** These are the cases where the model
correctly predicted the absence of malignancy (benign tumors).

**False Positives (FP) - 5:** These are the cases where the model
incorrectly predicted malignancy (the tumors were actually benign).

**False Negatives (FN) - 14:** These are the cases where the model
incorrectly predicted the tumors as benign, but they were actually
malignant. This is a critical error in medical diagnosis, as it means
missing out on identifying actual cases of cancer.

**True Positives (TP) - 58:** These are the cases where the model
correctly identified malignant tumors.

#### Accuracy

The accuracy of your model is (132 + 58)/ (132+5+14+58) ≈ 0.9091 or
90.91%. This means that about 91% of the time, this model correctly
predicts the class of the tumor.
